### 2020-12-19 paper/sex_selection_spacing-ver5.tex
Specify that sex selection is inferred

% Mortality
There has been a substantial reduction in infant mortality over time, and the size of the 
reductions is inversely related to the mother's education.
Hence, there is now little difference in mortality risk across education groups if the 
birth took place more than 21 months from the prior birth.
Short birth spacing is still associated with higher mortality, although the effect is 
small for the best-educated women.
There is no evidence that repeated abortions are associated with higher infant mortality
for the child eventually born; boys born after long birth intervals to families with
well-educated mothers, only daughters, and high predicted sex ratios, the combination of
which suggests repeated sex-selective abortions, have no higher infant mortality than other
children.

% Implications 
The results here paint a less rosy picture of India's prospects for a continued reduction 
in population growth than generally accepted. 
With predicted cohort fertility still substantially higher than the fertility rate, 
India's total fertility rate will likely stabilize or even increase as birth intervals slow 
their lengthening. 
The more successful the attempts at combatting sex selection are, the more likely an 
increase in the total fertility rate will be. 
Furthermore, the rapid decline in infant mortality risk, combined with likely future 
declines as the proportion of very short birth intervals falls, may also slow the 
reduction in population growth.

@@

There are two critical questions that future research should address.
First, sex selection means that girls-only families are less likely to have
very short birth intervals, which may reduce sibling competition. 
Hence, better health outcomes for girls with sex selection could be an unintended 
side-effect, rather than the result of girls becoming more valued as is often 
assumed \citep{Hu2015}. 
Comparing prior children's outcomes across sex composition and the sex of the next child 
could be a way to understand why girls' health outcomes improve in the presence of sex selection.

Second, what is the relationship between female labor force participation and sex selection? 
Women may be staying out of the labor market precisely because sex selection makes 
them more likely to have a boy and increases the expected birth spacing. 
Better job opportunities for women would affect sex selection for two reasons. 
First, it makes it more expensive to be out of the labor market for long periods. 
Second, it would moderate the differential in potential earnings between husband and wife 
and make it more attractive to invest in daughters' human capital. 
This approach could, however, be a double-edged sword. 
If better job opportunities further lower fertility, the use of sex selection may increase, 
everything else being equal.
Understanding the trade-off between long-term benefits from improvements in women's labor 
force participation and short-term costs from potential increases in sex selection is of 
paramount importance.

>>>>

% Mortality
There has been a substantial reduction in infant mortality over time, and the size of the 
reductions is inversely related to the mother's education.
Hence, there is now little difference in mortality risk across education groups if the 
birth took place more than 21 months from the prior birth.
Short birth spacing is still associated with higher mortality, although the effect is 
small for the best-educated women.
There is no evidence that repeated abortions are associated with higher infant mortality
for the child eventually born; boys born after long birth intervals to families with
well-educated mothers, only daughters, and high predicted sex ratios---the combination of
which suggests repeated sex-selective abortions---have no higher infant mortality than other
children.

% Implications 
The results here paint a less rosy picture of India's prospects for a continued reduction 
in population growth than generally accepted. 
With predicted cohort fertility still substantially higher than the period fertility rate, 
India's total fertility rate will likely stabilize or even increase as birth intervals slow 
their lengthening. 
The more successful the attempts at combatting sex selection are, the more likely an 
increase in the total fertility rate will be. 
Furthermore, the rapid decline in infant mortality risk, combined with likely future 
declines as the proportion of very short birth intervals falls, may also slow the 
reduction in population growth.

@@

There are two critical questions that future research should address.
First, access to sex selection means that girls-only families are less likely to have
very short birth intervals, which may reduce sibling competition. 
Hence, better health outcomes for girls with sex selection is available could be an 
unintended side-effect, rather than the result of girls becoming more valued as is often 
assumed \citep{Hu2015}. 
Comparing prior children's outcomes across sex composition and the sex of the next child 
could be a way to understand why girls' health outcomes improve in the presence of sex selection.

Second, what is the relationship between female labor force participation and sex selection? 
Women may be staying out of the labor market precisely because sex selection makes 
them more likely to have a boy and increases the expected birth spacing. 
Better job opportunities for women could reduce sex selection use for two reasons. 
First, it becomes more expensive to be out of the labor market for long periods. 
Second, the differential in potential earnings between husband and wife would decline, 
making it more attractive to invest in daughters' human capital.  
This approach could, however, be a double-edged sword. 
If better job opportunities further lower fertility, the use of sex selection may increase, 
everything else being equal.
Understanding the trade-off between long-term benefits from improvements in women's labor 
force participation and short-term costs from potential increases in sex selection is of 
paramount importance.
